Adaptive Scheduling of Scientific Workflows

### 2024-04-29 main.tex
first draft draft for the dynamic scenario

    \begin{table}[h]
        \begin{center}
            \begin{tabular}{rl}
                \hline
                \textbf{Symbol} & \textbf{Meaning}  \\
                \hline
                $G = (V, E)$  & Workflow graph, set of vertices (tasks) and edges  \\
                $\parents{u}$, $\children{u}$ & Parents of a task $u$, children of a task $u$ \\
                $m_u$& Memory weight of task $u$ \\
                $w_u$   & Workload of  task $u$  (normalized execution time)    \\
                $c_{u,v}$   & Communication volume along the edge $(u,v)\in E$ \\
                $F$, $\mathcal{F}$ & A partitioning function and the partition it creates \\
                $V_i$ & Block number $i$\\ %\wrt~some $F$   \\
                $\cluster$, $k$   & Computing system and its number of processors   \\
                $p_j$, proc($V_i$)  & Processor number $j$, processor of block $V_i$ \\
                $M_j$, $s_j$   & Memory size and speed of processor $p_j$   \\
                $\beta$ & Bandwidth in the compute system  \\
                $\bottomlevel{u}$   & Bottom weight of task $u$   \\
                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$ \\
                $\Gamma = (\mathcal{V}, \mathcal{E})$ & Quotient graph, its vertices and its edges  \\
                $r_u$, $r_{V_i}$  & Memory requirement of  task $u$ and of block $V_i$   \\
                $r_{\max}$   & Maximum memory requirement in a workflow   \\
                $\criticalpath$ & Critical path in a workflow  \\
                \hline
            \end{tabular}
        \end{center}
        \caption{Notation.} \label{tabnotation}
    \end{table}

@@

@@

    \begin{algorithm}[h!]
        \caption{Restore Validity}
        \label{alg:RestoreValidity}
        \begin{algorithmic}[1]
            \Procedure{Restore Validity}{$\Gamma$, $W$ }\\
            \Comment{Input: initial schedule as quotient tree $\Gamma$, new  weights $W$}
            \State $\Gamma \gets $ \textsc{QuotientDAG}($\Gamma$, $W$ ); \Comment Quotient DAG with new weights

@@

            \For{$\nu_i \in D$}
                \Comment Restore the validity of each damaged block
                \While{$r_{\nu_i} > M_{proc(\nu_i)}$}
                    \State $MS_{min}\gets \infty; v_{min} \gets NULL; \nu_{min} \gets NULL$
                    \Comment Moving $v_{min}$ to $\nu_{min}$ gives $MS_{min}$
                    \For{$v \in \nu_i$}
                        \For{$u \in ( \Pi_v \cup C_v) \cap (\Pi_{\nu_i} \cup C_{\nu_i})$}   \label{line:neighb}
                        \State $\Gamma' \gets \Gamma \mid \nu(v) == \nu(u)$
                        \If{ $r_{\nu(u)} > M_{proc(\nu(u))}$ and \textsc{Makespan}($\Gamma'$)$\leq MS_{min}$}\label{line:makespan}
                        \State $MS_{min} \gets $\textsc{Makespan}($\Gamma'$); $v_{min} \gets v$; $\nu_{min} \gets \nu(u);$
                        \EndIf
                        \EndFor
                    \EndFor
                    \State  $\Gamma \gets \Gamma \mid \nu(v_{min}) == \nu_{min}$ \label{line:movebest}
                \EndWhile
            \EndFor
            \State return $\Gamma$
            \EndProcedure
        \end{algorithmic}
    \end{algorithm}

@@

    \begin{algorithm}[h!]
        \caption{Repartition and Merge}
        \label{alg:RepartitionMerge}
        \begin{algorithmic}[1]
            \Procedure{Repartition and Merge}{$\Gamma$, $\cluster$ }\\
            \Comment{Input: quotient tree $\Gamma$, execution environment $\cluster$}
            \State $CP \gets $\textsc{Criticalpath}($\Gamma$) \label{line:cp}
            \State $Free \gets \{ p \in \cluster \mid p$ is free$\}$
            \While{$\mid Free \mid \neq 0$}\label{line:whilefree}
            \For{$\nu \in CP$}
                \State $\{ \nu_1, \nu_2,\dots\}\gets$\textsc{Partition}($\nu$,2) \label{line:repartition}
                \For{$\nu' \in \{ \nu_1, \nu_2,\dots\}$}
                    \If{$r_\nu' \leq Free.$head()}
                        \State proc($\nu'$)$\gets Free.$head();
                    \Else \label{line:ifnotfits}
                    \State \textsc{Merge}$\nu' \in \{ \nu_1, \nu_2,\dots\}$; break;
                    \Comment Merge the parts back and leave this partitioning, as it does not fit
                    \EndIf
                \EndFor
            \EndFor
            \EndWhile

>>>>

    \begin{table}[h]
        \begin{center}
            \begin{tabular}{rl}
                \hline
                \textbf{Symbol}                       & \textbf{Meaning}                                         \\
                \hline
                $G = (V, E)$                          & Workflow graph, set of vertices (tasks) and edges        \\
                $\parents{u}$, $\children{u}$         & Parents of a task $u$, children of a task $u$            \\
                $m_u$                                 & Memory weight of task $u$                                \\
                $w_u$                                 & Workload of task $u$  (normalized execution time)          \\
                $c_{u,v}$                             & Communication volume along the edge $(u,v)\in E$         \\
                $F$, $\mathcal{F}$                    & A partitioning function and the partition it creates     \\
                $V_i$                                 & Block number $i$                                         \\ %\wrt~some $F$   \\
                $\cluster$, $k$                    & Computing system and its number of processors           \\
                $p_j$, proc($V_i$)                          & Processor number $j$, processor of block $V_i$                 \\
                $M_j$, $s_j$                               & Memory size and speed of processor $p_j$                          \\
                $\beta$                     & Bandwidth in the compute system                                \\
                $\bottomlevel{u}$                      & Bottom weight of task $u$ \\
                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$               \\
                $\Gamma = (\mathcal{V}, \mathcal{E})$                      & Quotient graph, its vertices and its edges        \\
                $r_u$, $r_{V_i}$                            & Memory requirement of task $u$ and of block $V_i$                 \\
                $r_{\max}$                       & Maximum memory requirement in a workflow                              \\
                $\criticalpath$                       & Critical path in a workflow                              \\
                \hline
            \end{tabular}
        \end{center}
        \caption{Notation.} \label{tabnotation}
    \end{table}

@@

    \subsection{Dynamic Scenario}

    In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
    This information may include, memory usage, runtime, graph structures, or the status of the underlying infrastructure.
    In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
    By incorporating dynamic monitoring values, e.g., the resources a task consumed, the runtime environment can incorporate the data into the prediction model to provide more accurate resource predictions.
    Also the underlying infrastructure can change during the workflow execution.
    Examples are processor failures, node recoveries, or acquisition of new nodes.
    However, also when the hardware of the infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
    As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during the workflow runtime, the previous schedule becomes invalid and a new one must be calculated.

    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures of approximately 15\% is shown~\cite{}.
    Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
    %For instance, Nadeen~et~al.\cite{} report an error of 10\%, 11\%, and 15\% while the task prediction errors shows a normal and exponential distribution.
    %Bader~et~al.~ report a prediction error between 13\% and 17\% for their method, showing an exponential task error distribution.
    % @Svetlana, willst du sowas für deine Experimente? Also die Daten, welche du dann konfigurieren kannst?
    Such a dynamic execution environment necessitates for a dynamic scheduling method where the schedule can be recomputed during the workflow execution.

@@

    \begin{algorithm}[h!]
        \caption{Restore Validity}
        \label{alg:RestoreValidity}
        \begin{algorithmic}[1]
            \Procedure{Restore Validity}{$\Gamma$, $W$ }
                \\
                \Comment{Input: initial schedule as quotient tree $\Gamma$, new weights $W$}
                \State $\Gamma \gets $ \textsc{QuotientDAG}($\Gamma$, $W$ ); \Comment Quotient DAG with new weights

@@

                \For{$\nu_i \in D$}
                    \Comment Restore the validity of each damaged block
                    \While{$r_{\nu_i} > M_{proc(\nu_i)}$}
                        \State $MS_{min}\gets \infty; v_{min} \gets NULL; \nu_{min} \gets NULL$
                        \Comment Moving $v_{min}$ to $\nu_{min}$ gives $MS_{min}$
                        \For{$v \in \nu_i$}
                            \For{$u \in ( \Pi_v \cup C_v) \cap (\Pi_{\nu_i} \cup C_{\nu_i})$}
                                \label{line:neighb}
                                \State $\Gamma' \gets \Gamma \mid \nu(v) == \nu(u)$
                                \If{ $r_{\nu(u)} > M_{proc(\nu(u))}$ and \textsc{Makespan}($\Gamma'$)$\leq MS_{min}$}
                                    \label{line:makespan}
                                    \State $MS_{min} \gets $\textsc{Makespan}($\Gamma'$); $v_{min} \gets v$; $\nu_{min} \gets \nu(u);$
                                \EndIf
                            \EndFor
                        \EndFor
                        \State  $\Gamma \gets \Gamma \mid \nu(v_{min}) == \nu_{min}$ \label{line:movebest}
                    \EndWhile
                \EndFor
                \State return $\Gamma$
            \EndProcedure
        \end{algorithmic}
    \end{algorithm}

@@

    \begin{algorithm}[h!]
        \caption{Repartition and Merge}
        \label{alg:RepartitionMerge}
        \begin{algorithmic}[1]
            \Procedure{Repartition and Merge}{$\Gamma$, $\cluster$ }
                \\
                \Comment{Input: quotient tree $\Gamma$, execution environment $\cluster$}
                \State $CP \gets $\textsc{Criticalpath}($\Gamma$) \label{line:cp}
                \State $Free \gets \{ p \in \cluster \mid p$ is free$\}$
                \While{$\mid Free \mid \neq 0$}
                    \label{line:whilefree}
                    \For{$\nu \in CP$}
                        \State $\{ \nu_1, \nu_2,\dots\}\gets$\textsc{Partition}($\nu$,2) \label{line:repartition}
                        \For{$\nu' \in \{ \nu_1, \nu_2,\dots\}$}
                            \If{$r_\nu' \leq Free.$head()}
                                \State proc($\nu'$)$\gets Free.$head();
                            \Else
                                \label{line:ifnotfits}
                                \State \textsc{Merge}$\nu' \in \{ \nu_1, \nu_2,\dots\}$; break;
                                \Comment Merge the parts back and leave this partitioning, as it does not fit
                            \EndIf
                        \EndFor
                    \EndFor
                \EndWhile
